---
title: "Emmy awards"
date: 2021-09-22
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-09-21")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

nominees <- tt$nominees %>% 
  mutate(
    category = category %>% 
      str_remove("\\s-\\s\\d{4}$") %>% 
      str_to_lower()
  )

```

Need to regex for "category"

```{r}

nominees %>% 
  count(category, sort = TRUE)

```

Lots of NAs in "production"

```{r}

nominees %>% 
  count(production, sort = TRUE) 

```

Winners are 20 percent of all nominees

```{r}

nominees %>% 
  count(type, sort = TRUE) %>% 
  mutate(prop = n / sum(n))

```

Famous titles

```{r}

nominees %>% 
  count(title, sort = TRUE)

```

More nominees in recent years

```{r}

nominees %>% 
  count(year, sort = TRUE)

range(nominees$year)

nominees %>% 
  ggplot(aes(year)) +
  geom_bar() +
  labs(x = NULL, y = "# of nominees",
       title = "# of nominees have significantly increased since 2007",
       caption = "Source: emmys.com")

```

Distributors expand from Big 3 networks to Cable, and then Internet

```{r}

nominees %>% 
  count(distributor, sort = TRUE)

nominees %>% 
  group_by(distributor) %>% 
  slice_min(year, n = 1) %>% 
  distinct(distributor, year) %>% 
  ggplot(aes(year)) +
  geom_bar() +
  labs(x = NULL, y = "# of distributors debuted",
       title = "# of distributors have also increased since 2007",
       caption = "Source: emmys.com"
       )

```

```{r}

p1 <- nominees %>% 
  filter(!is.na(distributor)) %>% 
  group_by(distributor) %>% 
  mutate(n_years = length(unique(year))) %>%
  ungroup() %>% 
  filter(n_years > 4) %>% 
  distinct(year, distributor) %>% 
  mutate(
    distributor = distributor %>% 
      fct_reorder(year, .fun = min) %>% 
      fct_rev()
    ) %>% 
  ggplot(aes(year, distributor)) +
  geom_point(size = .5) +
  labs(x = NULL, y = NULL,
       title = "Transmission media expand from\nradio wave to cable, satellite,\nand the internet",
       subtitle = "Emmy nominee distributors\nwith more than 4 year nominations",
       caption = "Source: emmys.com")

ggsave("image/emmy_disties.png", p1, width = 4, height = 8)

```
