---
title: "NBER papers"
date: 2021-09-28
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

papers <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')

authors <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')

programs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')

paper_authors <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')

paper_programs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')


```


# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

joined_df <- left_join(papers, paper_authors) %>% 
  left_join(authors) %>% 
  left_join(paper_programs) %>% 
  left_join(programs)%>% 
  mutate(
    catalogue_group = str_sub(paper, 1, 1),
    catalogue_group = case_when(
      catalogue_group == "h" ~ "Historical",
      catalogue_group == "t" ~ "Technical",
      catalogue_group == "w" ~ "General"
    ),
    .after = paper
  ) 

joined_df

```


# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

by_paper <- joined_df %>% 
  distinct(paper, catalogue_group, year, month, title,
           program, program_desc, program_category)
  
by_paper %>% 
  count(catalogue_group, sort = TRUE)

by_paper %>% 
  count(program, program_desc, sort = TRUE)

by_paper %>% 
  count(program_category, sort = TRUE)

```



```{r}

by_paper %>% 
  count(year, program_category) %>% 
  complete(year, program_category, fill = list(n = 0)) %>% 
  mutate(
    program_category = if_else(is.na(program_category), "N.A.", program_category),
    program_category = fct_relevel(program_category, "N.A.")
    ) %>% 
  ggplot(aes(year, n, fill = program_category)) +
  geom_area(position = "fill") +
  scale_y_continuous(name = "Share in # of NBER papers", labels = scales::percent, expand = c(0, 0)) +
  scale_x_continuous(name = "Year", expand = c(0, 0)) +
  labs(fill = "Program category",
       title = "'Micro' dominates even after the financial crisis",
       subtitle = "Lucus critique, and then RCT may have contributed",
       caption = "Source: NBER")

```


```{r}

library(tidytext)

p <- by_paper %>% 
  mutate(decade = 10 * (year %/% 10)) %>% 
  add_count(decade) %>% 
  mutate(decade = glue::glue("{decade}s ({n})")) %>% 
  select(decade, title) %>% 
  unnest_tokens(word, title) %>% 
  count(decade, word) %>% 
  bind_tf_idf(word, decade, n) %>% 
  group_by(decade) %>% 
  slice_max(tf_idf, n = 10) %>% 
  ungroup() %>% 
  mutate(word = reorder_within(word, tf_idf, decade)) %>% 
  ggplot(aes(tf_idf, word)) +
  geom_col(aes(fill = decade), show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(vars(decade), scales = "free_y") +
  labs(x = "Term frequency times inverse document frequency in title",
       y = "Top 10 words",
       title = "2020s is by far the most peculiar decade",
       subtitle = "( ) denotes # of NBER papers",
       caption = "Source: NBER")

p

```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/nber_tf_idf.png", p, width = 8, height = 4)

```
