---
title: "Pride Corporate Accountability Project"
date: 2022-06-07
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

library(scales)
theme_set(theme_light())

library(janitor)
library(lubridate)
library(ggrepel)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-06-07")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
contribution_data_all_states <- tt$contribution_data_all_states %>% 
  clean_names()
  
corp_by_politician <- tt$corp_by_politician %>% 
  clean_names()

donors <- tt$donors %>% 
  clean_names()

fortune_agg <- tt$fortune_agg %>% 
  clean_names()

fortune_aggregates <- tt$fortune_aggregates %>% 
  clean_names()

pride_aggregates <- tt$pride_aggregates %>% 
  clean_names()

pride_sponsors <- tt$pride_sponsors %>% 
  clean_names()

static_list <- tt$static_list %>% 
  clean_names()

```

```{r}
rm(fortune_agg)

contribution_data_all_states %>% 
  count(state, sort = TRUE)

contribution_data_all_states %>% 
  summarize(total_amount = sum(amount, na.rm = TRUE)) # Total 27M$

contribution_data_all_states %>% 
  group_by(politician) %>% 
  summarize(amount_by_politician = sum(amount, na.rm = TRUE)) %>% 
  arrange(desc(amount_by_politician)) # 109 politicians

range(contribution_data_all_states$date, na.rm = TRUE) # from 2013 Jan to 2022 May

contribution_data_all_states %>% 
  count(pride_and_sponsor_match, sort = TRUE)

contribution_data_all_states %>% 
  filter(!is.na(pride_and_sponsor_match)) %>% 
  group_by(politician) %>% 
  summarize(amount_by_politician = sum(amount, na.rm = TRUE)) %>% 
  arrange(desc(amount_by_politician)) # 103 politicians

# roughly match corp_by_politician with # total 4M$, 102 politicians
corp_by_politician %>% 
  count(state, sort = TRUE) 

contribution_data_all_states %>% 
  count(donor_type, wt = amount, sort = TRUE)

range(pride_sponsors$year) # only 2021 and 2022

fortune_aggregates %>% 
  anti_join(pride_sponsors, by = "company") # FedEx and Fed Ex

fortune_aggregates %>% 
  anti_join(static_list, by = "company") # all 117 are included in 126 companies
# static_list 126 companies look like all Fortune 500 companies

contribution_data_all_states %>% 
  filter(pride | hrc_business_pledge) %>% 
  summarize(amount = sum(amount, na.rm = TRUE)) # 2.2M$
```

```{r}
corp_by_politician %>% 
  filter(title == "Governor")

```

# Visualize

Using your processed dataset, create your unique visualization.

Should I concentrate on years from 2019 to 2021?

```{r Visualize}
contribution_data_all_states %>% 
  filter(pride | hrc_business_pledge) %>% 
  group_by(year = year(date), state) %>% 
  summarize(amount = sum(amount, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(factor(year), amount, fill = state)) +
  geom_col() +
  scale_y_continuous(labels = dollar) +
  labs(title = "Betrayal suddently increased in 2019?",
    subtitle = "Donation betraying either Pride or HRC Business Pledge",
       x = NULL, y = NULL, fill = "State",
       caption = "Source: Data for Progress")
```

```{r}
contribution_data_all_states %>% 
  filter(pride, !hrc_business_pledge) %>% 
  group_by(company) %>% 
  summarize(amount = sum(amount, na.rm = TRUE)) %>% 
  arrange(desc(amount))
# pride only 28, hrc_business_pledge only 8, both 7 companies

```

Greg Abbott, Texas Governor

```{r}
p <- contribution_data_all_states %>% 
  mutate(
    fortune = if_else(pride_and_sponsor_match %in% c("Both", "Fortune"),
    "Fortune 500 companies", "Non Fortune 500"),
    donor_name = str_to_title(donor_name)
    ) %>% 
  filter(politician == "Greg Abbott",
         year(date) == 2021) %>%
  group_by(fortune, donor_name) %>% 
  summarize(amount = sum(amount, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(amount)) +
  geom_histogram(aes(fill = fortune)) +
  geom_text_repel(aes(x = amount, y = 30, label = donor_name), size = 3,
                  data = . %>% filter(amount > 1e5)) +
  geom_text_repel(aes(x = amount, y = 30, label = donor_name), size = 3,
                  data = . %>% filter(amount >= 1e5, fortune == "Fortune 500 companies")) +
  scale_x_log10(labels = dollar) +
  guides(fill = "none") +
  facet_wrap(vars(fortune), ncol = 1) +
  labs(x = "Amount", y = "# of donors",
       title = "Donations to Texas Governor, Greg Abbott in 2021",
       subtitle = "Amount distribution and some largest donors",
       caption = "Source: Data for Progress")

p
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/pride_corp_account.png", p, width = 6, height = 4)

```
