---
title: "Company reputation poll"
date: 2022-05-31
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-05-31")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
poll <- tt$poll
reputation <- tt$reputation
```

```{r}
nrow(reputation) # 700 rows

reputation %>% 
  distinct(name) # 7 names

reputation %>% 
  count(company, sort = TRUE) # 100 companies

reputation %>% 
  group_by(company) %>% 
  summarize(mean = mean(score)) %>% 
  filter(company == "3M Company")

nrow(poll) # 500 rows

poll %>% 
  filter(company == "3M Company")

poll %>% 
  distinct(company) # 100 companies

poll %>% 
  distinct(year) # basically 5 years

poll %>% 
  filter(is.na(year))

poll %>% 
  filter(company == "Shein")

poll <- poll %>% 
  mutate(year = if_else(is.na(year), 2018, year))
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}

lvl <- reputation %>% 
  count(industry, sort = TRUE) %>%  # 19 industries
  pull(industry)

reputation %>% 
  mutate(
    name = factor(name,
                       levels = c("TRUST", "ETHICS", "GROWTH",
                                  "P&S", "CITIZENSHIP", "VISION", "CULTURE")),
    industry = factor(industry, levels = lvl)
    ) %>% 
  ggplot(aes(name, score)) +
  geom_line(aes(group = company, color = company)) +
  facet_wrap(vars(industry)) +
  guides(color = "none") +
  labs(x = NULL) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}

p <- reputation %>% 
  group_by(name) %>% 
  summarize(
    median = median(score),
    ll = quantile(score, prob = 0.1),
    ul = quantile(score, prob = 0.9)
  ) %>% 
  mutate(name = fct_reorder(name, median)) %>% 
  ggplot(aes(median, name)) +
  geom_pointrange(aes(xmin = ll, xmax = ul)) +
  labs(x = "Score", y = "Attributes",
       title = "Scoring is strict on CITIZENSHIP,\nand lax on VISION, P&S and GROWTH",
       subtitle = "median point, and range of 80 out of 100 companies",
       caption = "Source: Axios and Harris Poll")

p
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/company-reputation.png", p, width = 6, height = 4)

```
