---
title: "Chocolate bar ratings"
date: 2022-01-18
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

library(scales)
theme_set(theme_light())

```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-01-18")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

chocolate <- tt$chocolate %>% 
  mutate(
    cocoa_percent = cocoa_percent %>% 
      str_sub(end = -2L) %>% 
      as.numeric()
  ) %>% 
  separate(ingredients, 
           into = c("n_ingredients", "ingredients"),
           sep = "-") %>% 
  mutate(n_ingredients = as.numeric(n_ingredients)) %>% # max is 6
  mutate(ingredients = str_trim(ingredients))

```


# Visualize

ref is not unique. ref shows how new entry is.
ref up to 600 tends to have lower rating with more variance. 

```{r Visualize}

dim(chocolate) # 2530 rows

chocolate %>% 
  count(ref, sort = TRUE)

range(chocolate$ref)

chocolate %>% 
  ggplot(aes(ref)) +
  geom_histogram(boundary = 0, binwidth = 100)

summarize_rating <- function(df) {
  df %>% 
    summarize(
    n = n(),
    avg_rating = mean(rating),
    median_rating = median(rating)
    )
}

chocolate %>% 
  mutate(ref = cut(ref, breaks = 0:14 * 200)) %>% 
  group_by(ref) %>% 
  summarize_rating()

library(ggridges)

draw_density <- function(df, var) {
  df %>% 
  ggplot(aes(rating, {{ var }}, fill = {{ var }})) +
  geom_density_ridges(show.legend = FALSE)
}

chocolate %>% 
  mutate(ref = cut(ref, breaks = 0:14 * 200)) %>% 
  draw_density(ref)
  
```

review_date up to 2009 tends to be lower.

```{r}

range(chocolate$review_date)

chocolate %>% 
  ggplot(aes(rating, factor(review_date))) +
  geom_boxplot()

```

Belgium tends to be lower.

```{r}

chocolate %>% 
  count(company_manufacturer, sort = TRUE)

chocolate %>% 
  count(company_location, sort = TRUE)

chocolate %>% 
  mutate(
    company_location = company_location %>% 
      fct_lump_n(10) %>% 
      fct_reorder(rating)
    ) %>% 
  draw_density(company_location)

```

Blend bean tends to be lower.

```{r}

chocolate %>% 
  count(country_of_bean_origin, sort = TRUE)

chocolate %>% 
  mutate(
    country_of_bean_origin = country_of_bean_origin %>% 
      fct_lump_n(10) %>% 
      fct_reorder(rating)
    ) %>% 
  draw_density(country_of_bean_origin)

```

cocoa_percent from 60 to 80 is most common and tends to have higher rating.

```{r}

chocolate %>% 
  ggplot(aes(cocoa_percent, rating)) +
  geom_point(alpha = 0.2)

range(chocolate$cocoa_percent)

chocolate %>% 
  ggplot(aes(cocoa_percent)) +
  geom_histogram(boundary = 40, binwidth = 5)

chocolate %>% 
  mutate(cocoa_percent = cut(cocoa_percent, breaks = 8:20 * 5)) %>% 
  group_by(cocoa_percent) %>% 
  summarize_rating()

chocolate %>% 
  mutate(cocoa_percent = cut(cocoa_percent, breaks = 10:20 * 5)) %>% 
  filter(!is.na(cocoa_percent)) %>% 
  draw_density(cocoa_percent)

```

n_ingredients from 2 to 3 is most common and tends to have higher rating.

```{r}

chocolate %>% 
  group_by(n_ingredients) %>% 
  summarize_rating()

chocolate %>% 
  ggplot(aes(factor(n_ingredients))) +
  geom_bar()

chocolate %>% 
  ggplot(aes(factor(n_ingredients), rating)) +
  geom_boxplot()

```

Beans, Sugar and Cocoa Butter are most common ingredients and tend to have high rating.

```{r}

chocolate %>% 
  separate_rows(ingredients, sep = ",") %>% 
  group_by(ingredients) %>% 
  summarize_rating()

p <- chocolate %>% 
  filter(!is.na(ingredients)) %>% 
  separate_rows(ingredients, sep = ",") %>% 
  add_count(ingredients) %>% 
  mutate(
    ingredients = recode(ingredients, "B" = "Beans", "S" = "Sugar", 
                         "S*" = "Sweetener other than\nwhite cane or beet sugar",
                         "C" = "Cocoa Butter",
                         "V" = "Vanilla", "L" = "Lecithin", "Sa" = "Salt"),
    ingredients = glue::glue("{ingredients} ({n})"),
    ingredients = fct_reorder(ingredients, rating)
    ) %>% 
  draw_density(ingredients) +
  
  labs(title = "Vanilla, sweetener and salt tend to\nlower rating",
       subtitle = "( ) denotes the number of chocolate bars",
       caption = "Source: Flavors of Cacao")

p

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/chocolate_bar.png", p, width = 6, height = 4)

```

# Try sankey diagram

```{r}

library(ggsankey)

chocolate %>% 
  mutate(
    company_location = fct_lump_n(company_location, 5),
    country_of_bean_origin = fct_lump_n(country_of_bean_origin, 5),
    n_ingredients = factor(n_ingredients)
  ) %>% 
  make_long(company_location, country_of_bean_origin, n_ingredients) %>% 
  ggplot(aes(x, node = node, next_x = next_x, next_node = next_node,
             fill = factor(node), label = node)) +
  geom_sankey(show.legend = FALSE) +
  geom_sankey_label(show.legend = FALSE) +
  theme_sankey()

```

# Imitate Julia Silge

*This template offers an opinionated guide on how to structure a modeling analysis. Your individual modeling analysis may require you to add to, subtract from, or otherwise change this structure, but consider this a general framework to start from. If you want to learn more about using tidymodels, check out our [Getting Started](https://www.tidymodels.org/start/) guide.*


```{r}
library(tidymodels)

```


## Explore data

Exploratory data analysis (EDA) is an [important part of the modeling process](https://www.tmwr.org/software-modeling.html#model-phases).

```{r}
chocolate %>% 
  ggplot(aes(rating)) +
  geom_histogram()
```

```{r}
library(tidytext)

tidy_chocolate <- chocolate %>% 
  unnest_tokens(word, most_memorable_characteristics)

tidy_chocolate %>% 
  count(word, sort = TRUE)
```

```{r}

tidy_chocolate %>% 
  group_by(word) %>% 
  summarize(
    n = n(),
    rating = mean(rating)
  ) %>% 
  ggplot(aes(n, rating)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = mean(chocolate$rating), color = "blue") +
  geom_text(aes(label = word), check_overlap = TRUE,
            vjust = "top", hjust = "left") +
  scale_x_log10()
```


## Build models

Let's consider how to [spend our data budget](https://www.tmwr.org/splitting.html):

- create training and testing sets
- create resampling folds from the *training* set

```{r}
set.seed(123)
choco_split <- initial_split(chocolate, strata = rating)
choco_train <- training(choco_split)
choco_test <- testing(choco_split)

set.seed(234)
choco_folds <- vfold_cv(choco_train, strata = rating)
choco_folds
```

Let's make a recipe for feature engineering.

```{r}
library(textrecipes)

choco_rec <- recipe(rating ~ most_memorable_characteristics,
                    data = choco_train) %>% 
  step_tokenize(most_memorable_characteristics) %>% 
  step_tokenfilter(most_memorable_characteristics, max_tokens = 100) %>% 
  step_tfidf(most_memorable_characteristics)

prep(choco_rec) %>% bake(new_data = NULL)
```


Let's create a [**model specification**](https://www.tmwr.org/models.html) for each model we want to try:

```{r}
rf_spec <-
  rand_forest(trees = 500) %>%
  set_mode("regression")

rf_spec
```

```{r}
svm_spec <- svm_linear() %>% 
  set_mode("regression")

svm_spec
```


To set up your modeling code, consider using the [parsnip addin](https://parsnip.tidymodels.org/reference/parsnip_addin.html) or the [usemodels](https://usemodels.tidymodels.org/) package.

Now let's build a [**model workflow**](https://www.tmwr.org/workflows.html) combining each model specification with a data preprocessor:

```{r}
svm_wf    <- workflow(choco_rec, svm_spec)
ranger_wf <- workflow(choco_rec, rf_spec)
```

If your feature engineering needs are more complex than provided by a formula like `sex ~ .`, use a [recipe](https://www.tidymodels.org/start/recipes/). [Read more about feature engineering with recipes](https://www.tmwr.org/recipes.html) to learn how they work.


## Evaluate models

These models have no tuning parameters so we can evaluate them as they are. [Learn about tuning hyperparameters here.](https://www.tidymodels.org/start/tuning/)

```{r}
doParallel::registerDoParallel(cores = 5)
contrl_preds <- control_resamples(save_pred = TRUE)

svm_rs <- fit_resamples(
  svm_wf,
  resamples = choco_folds,
  control = contrl_preds
)

ranger_rs <- fit_resamples(
  ranger_wf,
  resamples = choco_folds,
  control = contrl_preds
)
```

How did these two models compare?

```{r}
collect_metrics(svm_rs)
collect_metrics(ranger_rs)
```

We can visualize these results using an ROC curve (or a confusion matrix via `conf_mat()`):

```{r}
bind_rows(
  collect_predictions(svm_rs) %>%
    mutate(mod = "svm"),
  collect_predictions(ranger_rs) %>%
    mutate(mod = "ranger")
) %>% 
  ggplot(aes(rating, .pred, color = id)) +
  geom_jitter() +
  geom_abline() +
  facet_wrap(vars(mod)) +
  coord_fixed()
```

These models perform very similarly, so perhaps we would choose the simpler, linear model. The function `last_fit()` *fits* one final time on the training data and *evaluates* on the testing data. This is the first time we have used the testing data.

```{r}
final_fitted <- last_fit(svm_wf, choco_split)
collect_metrics(final_fitted)  ## metrics evaluated on the *testing* data
```

This object contains a fitted workflow that we can use for prediction.

```{r}
final_wf <- extract_workflow(final_fitted)
predict(final_wf, choco_test[55,])
```

You can save this fitted `final_wf` object to use later with new data, for example with `readr::write_rds()`.

```{r}
final_wf %>% 
  tidy() %>% 
  mutate(term = str_remove(term, "tfidf_most_memorable_characteristics_")) %>% 
  filter(term != "Bias") %>% 
  group_by(estimate > 0) %>% 
  slice_max(abs(estimate), n = 10) %>% 
  ungroup() %>% 
  mutate(term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(estimate, term, fill = estimate > 0)) +
  geom_col(show.legend = FALSE)
```

