---
title: "US House Election Results"
date: 2023-11-09
execute: 
  echo: true
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r}
#| label: Load

tt <- tt_load("2023-11-07")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r}
#| label: Readme
#| eval: interactive()

tt

```

# Glimpse Data

Take an initial look at the format of the data available.

```{r}
#| label: Glimpse

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r}
#| label: Wrangle

house <- tt$house
```

```{r}
house |> 
  count(party, sort = TRUE)

house |> 
  count(stage, sort = TRUE)

house |> 
  count(state, sort = TRUE)

house |> 
  count(year, sort = TRUE)
```

```{r}
house |> 
  filter(
    state == "NEW YORK", stage == "GEN", !runoff,
    !is.na(party)
    ) |> 
  count(district, year, party, sort = TRUE)

house |> 
  filter(
    state == "NEW YORK", stage == "GEN", !runoff,
    !is.na(party)
    ) |> 
  filter(district == "005", year == 1998) |> View()
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r}
#| label: Visualize

house |> 
  filter(
    state == "NEW YORK", stage == "GEN", !runoff,
    party %in% c("DEMOCRAT", "REPUBLICAN")
    ) |> 
  mutate(
    vote_ratio = candidatevotes / totalvotes,
    district = fct_reorder(district, totalvotes)
  ) |> 
  ggplot(aes(year, vote_ratio)) +
  geom_line(aes(color = party)) +
  scale_color_manual(values = c("blue", "red")) +
  facet_wrap(vars(district))

house |> 
  filter(
    state == "NEW YORK", stage == "GEN", !runoff,
    party %in% c("DEMOCRAT", "REPUBLICAN")
    ) |> 
  distinct(year, district, totalvotes) |> 
  ggplot(aes(factor(year), totalvotes)) +
  geom_boxplot()
```

```{r}
winners <- house |> 
  filter(stage == "GEN", !runoff) |> 
  group_by(state, year, district, candidate) |> 
  summarize(candidatevotes = sum(candidatevotes), .groups = "drop_last") |> 
  slice_max(candidatevotes)

deadvote <- house |> 
  filter(stage == "GEN", !runoff) |> 
  semi_join(winners, join_by(year, state, district, candidate)) |> 
  group_by(year, state, district) |> 
  summarize(
    totalvotes = sum(totalvotes),
    candidatevotes = sum(candidatevotes),
    .groups = "drop"
  ) |> 
  inner_join(house |> 
               filter(party %in% c("DEMOCRAT", "REPUBLICAN")) |>
               select(year, state, district, candidate, party)
               , join_by(year, state, district)) |> 
  mutate(
    deadvotes = totalvotes - candidatevotes,
    deadvotes_ratio = deadvotes / totalvotes,
    party = paste0(str_to_title(party), " won")
  )

deadvote |> 
  group_by(year, party) |> 
  summarize(deadvotes_ratio = sum(deadvotes) / sum(totalvotes),
            .groups = "drop") |> 
  ggplot(aes(year, deadvotes_ratio)) +
  geom_line(aes(color = party)) +
  scale_color_manual(values = c("blue", "red"))
  
library(ggridges)

p <- deadvote |> 
  ggplot(aes(deadvotes_ratio, factor(year))) +
  geom_density_ridges(aes(fill = party), alpha = 0.5,
                      show.legend = FALSE) +
  geom_vline(xintercept = 0.5, lty = 2) +
  facet_wrap(vars(party)) +
  scale_fill_manual(values = c("blue", "red")) +
  labs(x = "Ratio of dead votes, i.e. the votes a winner didn't take",
       y = NULL,
       title = "Dead vote ratios were low before 1990,\nregardless which party won",
     caption = "Source: MIT Election Data and Science Lab") +
  theme(strip.background = element_blank(),
        strip.text = element_text(color = "black"))
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/us_house_election_results.png", p, width = 6, height = 5)

```
