---
title: "The Global Human Day"
date: 2023-09-12
execute: 
  echo: true
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r}
#| label: Load

tt <- tt_load("2023-09-12")
```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r}
#| label: Readme
#| eval: interactive()

tt

```

# Glimpse Data

Take an initial look at the format of the data available.

```{r}
#| label: Glimpse

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

year: 49 years from 1973 to 2022 except 1982

```{r}
#| label: Wrangle

all_countries <- tt$all_countries
country_regions <- tt$country_regions
global_human_day <- tt$global_human_day
global_economic_activity <- tt$global_economic_activity
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r}
#| label: Visualize

global_human_day |> 
  ggplot(aes(hoursPerDay, Subcategory)) +
  geom_col(alpha = 0.5) +
  geom_errorbar(aes(xmin = hoursPerDay - uncertainty,
                    xmax = hoursPerDay + uncertainty))
```

```{r}
all_countries |> 
  filter(country_iso3 == "JPN") |> 
  ggplot(aes(hoursPerDayCombined, Subcategory)) +
  geom_col(alpha = 0.5) +
  geom_errorbar(aes(xmin = hoursPerDayCombined - uncertaintyCombined,
                    xmax = hoursPerDayCombined + uncertaintyCombined))

subcategories <- all_countries |> 
  filter(country_iso3 == "JPN") |> 
  pull(Subcategory)
```

```{r}
p <- bind_rows(
  global_human_day |> 
    mutate(country = "Global average"),
  all_countries |> 
    filter(country_iso3 == "JPN") |> 
    select(Subcategory, hoursPerDay = hoursPerDayCombined,
           uncertainty = uncertaintyCombined) |> 
    mutate(country = "Japan")
) |> 
  mutate(Subcategory = factor(Subcategory, levels = subcategories)) |> 
  ggplot(aes(hoursPerDay, Subcategory)) +
  geom_col(aes(fill = country), alpha = 0.5,
           position = "dodge") +
  geom_errorbar(aes(xmin = hoursPerDay - uncertainty,
                    xmax = hoursPerDay + uncertainty,
                    color = country),
                position = "dodge") +
  scale_fill_manual(values = c("gray50", "red"),
                    guide = guide_legend(reverse = TRUE)) +
  scale_color_manual(values = c("gray50", "red"),
                    guide = guide_legend(reverse = TRUE)) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 1)),
                     breaks = 0:10) +
  labs(x = "Hours per day", y = NULL,
       fill = NULL, color = NULL,
       title = "Japanese spend less time for sleep and social, and \nspend more time for passive than global average",
       subtitle = "Error bar shows plus minus one standard error",
     caption = "Source: The Global Human Day dataset") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank())

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/global_human_day.png", p, width = 7, height = 7)

```
