---
title: "Roundabouts across the world"
date: 2025-12-16
execute: 
  echo: true
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(tidytuesdayR)

theme_set(theme_light())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r}
#| label: Load

tt <- tt_load("2025-12-16")

```

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r}
#| label: Readme
#| eval: interactive()

tt

```

# Glimpse Data

Take an initial look at the format of the data available.

```{r}
#| label: Glimpse

tt |> 
  map(glimpse)
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r}
#| label: Wrangle

roundabouts_clean <- tt$roundabouts_clean
```

```{r}
skimr::skim(roundabouts_clean)
```

```{r}
roundabouts_clean |> 
  count(country, sort = TRUE)

roundabouts_clean |> 
  filter(country == "Japan") |> View()

roundabouts_clean |> 
  count(approaches, sort = TRUE)

roundabouts_clean |> 
  filter(approaches == 1) |> View()

roundabouts_clean |> 
  count(lane_type, sort = TRUE)

roundabouts_clean |> 
  filter(lane_type == "Ring junction (3+ mini-roundabouts in ring)") |> View()
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r}
#| label: Visualize

library("cartographr")

# 出力用紙サイズを設定する:set_output_sizeコマンド
# サイズを指定:sizeオプション;"A0","A1","A2","A3","A4","A5","A6",
# "small_poster","medium_poster","large_poster",c(280,280)などと指定
# set_output_size(c(280,280)) 
set_output_size("A6")

roundabouts_in_japan <- roundabouts_clean |> 
  filter(country == "Japan") |> 
  mutate(address = str_remove(address, " \\(Japan\\)")) |> 
  unite(name, name:address, sep = ", ") |> 
  mutate(
    name = name |> 
      str_remove("\\(unknown\\), ") |> 
      str_wrap(width = 30)
  )

roundabouts_in_japan |> 
  count(county_area, sort = TRUE)

roundabouts_in_japan |> 
  count(state_region, sort = TRUE)

# OpenStreetMapからデータを取得:get_osmdataコマンド
# 中心の緯度を指定:latオプション
# 中心の経度を指定:lonオプション
# 取得領域幅の指定:x_distanceオプション;距離はメートル
# 取得領域高さの指定:y_distanceオプション;省略で幅と用紙サイズの縦横比に基づく
# 縦横比の指定:aspect_ratioオプション
# 処理メッセージの表示/非表示:quietオプション;TRUE/FALSE
# 取得したデータを保持する:keepオプション;TRUE/FALSE
roundabouts_maps <- roundabouts_in_japan |> 
  mutate(
    maps = map2(long, lat, \(long, lat) get_osmdata(lat = long, lon = lat, x_distance = 50, aspect_ratio = 1, quiet = FALSE, keep = FALSE))
  )

length(roundabouts_maps)

roundabouts_maps[[18]][[3]]
plot_map(roundabouts_maps[[18]][[3]], palette = "tropical")

# 取得データをプロット:plot_mapコマンド
# データを指定:osmオプション
# パレットを適応:paletteオプション
# "alphabet","arctic","autumn","autumn-muted","bw","desert","evening",
# "gray","iberia","imhof","lines","metropolitan","midnight","minimal",
# "swiss", "tropical"が設定可能
plots <- vector("list", length = nrow(roundabouts_in_japan))

draw_each_roundabout <- function(i) {
  plot_map(roundabouts_maps[[18]][[i]], palette = "tropical") +
  geom_text(aes(x = lat, y = long, label = name),
            data = roundabouts_in_japan |> slice(i),
            hjust = 0.5, vjust = 0.5, color = "black")
}

draw_each_roundabout(1)
draw_each_roundabout(2)
draw_each_roundabout(3) # error
draw_each_roundabout(4)
draw_each_roundabout(5)
draw_each_roundabout(6)
draw_each_roundabout(7)
draw_each_roundabout(8)

plots <- map(c(1:2, 4:30), draw_each_roundabout)

library(patchwork)

rows_app_gt_4 <- roundabouts_in_japan |> 
  mutate(row = row_number()) |> 
  filter(approaches > 4) |> 
  pull(row)

plots[[1]]
plots[[6]]
plots[[5]]

plots_pc <- plots[[1]]

for (i in rows_app_gt_4[2:5] - 1) {
  plots_pc <- plots_pc + plots[[i]]
}

p <- plots_pc +
  plot_layout(ncol = 3) +
  plot_annotation(
    title = "There are only 5 roundabouts with more than 4 approaches in Japan",
    subtitle = "Sakuragaoka, Tama City roundabout is the only one I personally have watched",
    caption = "Source: roundabouts R package",
    theme = theme(
      plot.title.position = "plot",
      plot.caption.position = "plot"
      )
  )
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave("image/roundabouts_across_the_world.png", p, width = 4, height = 3)
```
